The Impact of Extreme Weather Events on Property Insurance Pricing

### 2023-12-11 manuscript/manuscript.tex
more redundancy removed and modified for easy reading

\begin{abstract}
    In recent years, the increasing severity of weather disasters like hurricanes, wildfires, and hail storms and 
    their impact on homeowners and their finances. It explores how the frequency and severity of these catastrophes influence property 
    insurance prices. Using statistical analysis and data from reliable sources such as the Insurance Information Institute and Aon's 
    2023 Weather, Climate, and Catastrophe Insight Report, the study aims to provide practical insights. These insights aim to inform 
    both insurance industry practices and public policy decisions, focusing on improving resilience to such catastrophes and easing 
    the financial burden on homeowners.
\end{abstract}

\section{Introduction}
\label{sec:intro}
As the frequency and severity of weather catastrophe increased in recent years, it has been shown to pose threats to homeowners' 
safety and finance as well. The study explores the rising frequency and severity of recent weather disasters, threatening homeowners' 
safety and finances. It investigates their impact on property insurance pricing, emphasizing the catastrophic consequences for 
homeowners and insurers. Managing insurance premiums, deductibles, and coverage limitations becomes challenging for homeowners, 
while insurers face heightened financial risks.

Unlike prior research \cite{hurricaneeco}, our study delves deeper into the dynamics of frequency and severity in property insurance 
pricing, seeking effective ways to alleviate the financial burden during major disasters. Drawing data from reliable sources—such as 
the Insurance Information Agency, National Insurance Centers for Environmental Information, and Aon's 2023 Weather, Climate, and 
Disaster Investigation Report—we employ statistical analysis and data science techniques. The objective is to enhance homeowner 
well-being and ensure stability in the property insurance market amidst increasing weather-related risks.

The rest of the paper is organized as follows.
The data will be presented in Section~\ref{sec:data}.
The methods are described in Section~\ref{sec:meth}.
The results are reported in Section~\ref{sec:resu}.
A discussion concludes in Section~\ref{sec:disc}.

\section{Data}
\label{sec:data}
In order to understand the impact of weather catastrophe on property insurance pricing, we will first look at a dataset of 
billion-dollar disasters. This dataset contains important information about the frequency, financial cost, and some other parameters. 
Our data sources include the National Centers for Environmental Information (NCEI)\cite{ncei}, the Insurance Information Institute 
(III)\cite{iii}, Aon\cite{aon} and the National Association of Insurance Commissioners (NAIC)\cite{naic}, which are all known for their 
reliability in documenting catastrophic events and their associated economic and insured losses.

@@

The figure \ref{fig:disaster_threats} shows the cumulative billion-dollar disaster costs in the United States from 1980-2023 on a year-to-date basis. It 
illustrates how annual disaster costs accumulate through the year, with key major disaster event years highlighted. The table 
provides context on total disaster costs over time.

@@

The figure \ref{fig:regional_disasters} shows the number of billion-dollar disaster events by type in the United States from 1980-2023. It breaks down event counts 
for droughts, flooding, freezes, severe storms, tropical cyclones, wildfires, and winter storms. The total disaster cost is also shown 
accumulated across years. The chart gives overview insight into disaster frequency and cost by peril.

@@

The chart \ref{fig:10economic_loss} presents the ten most expensive global insured catastrophe losses from 1900 to 2022, highlighting 
the rising risks and costs that property insurers are facing. The chart highlights the insurance industry's climate-change costs, which 
require immediate adjustments to rates, coverages, and risk-mitigation efforts, with a focus on vulnerable coastal and island regions.

@@

The chart \ref{fig:10global_loss} showing the top ten economically destructive tropical cyclones worldwide from 1900 to 2022 
highlights the increasing risks posed by hurricanes in a changing climate. In particular, the United States has been hit by all ten 
of the costliest storms since 2000, suggesting a concerning trend of recurring severe hurricane seasons for coastal regions.

@@

\begin{equation}
    \label{eq:freq}
    \text{Frequency} = \frac{\text{Number of Loss Events}}{\text{Exposure Units}}
\end{equation}

The frequency equation \ref{eq:freq} calculates the frequency of loss events by dividing the number of loss events by exposure units.

\begin{equation}
    \label{eq:sev}
    \text{Severity} = \frac{\text{Total Loss Amount}}{\text{Number of Loss Events}}
\end{equation}

The severity equation \ref{eq:sev} is determined by dividing the total loss amount by the number of loss events.

\begin{equation}
    \label{eq:losscost}
    \text{Loss Cost} = \text{Frequency} \times \text{Exposure Units}
\end{equation}

The loss count equation \ref{eq:losscost} is the product of frequency and exposure units, providing insights into the overall loss count due to catastrophic events.

\begin{equation}
    \label{eq:premium}
    \text{Premium} = \lambda \cdot \text{Severity} \cdot \text{Exposure Units}
\end{equation}

The premium equation \ref{eq:premium} calculates the estimated premium based on the frequency, severity, and exposure units, which 
shows the financial consequences for both homeowners and insurance companies.

By understanding these fundamental equations, we hope to gain a thorough understanding of the relationship between weather-related disasters 
and property insurance pricing. Like mentioned earlier, these equations will serve as a basis for our data analysis, allowing us to 
effectively assess the impact of catastrophic events on property insurance while also assisting us in meeting our research objectives.

>>>>

\begin{abstract}
    In recent years, with frequency and severity of weather catastrophes both increasing, it has been shown to make a significant impact 
    on homeowners as well as making a financial burden. We will be exploring how the frequency and severity of these catastrophes 
    influence property insurance prices through statistical analysis and data from reliable sources, such as the Insurance Information 
    Institute and Aon's 2023 Weather, Climate, and Catastrophe Insight Report. The objective of this study aims to provide practical 
    insights to inform both insurance industry practices and public policy decisions, in addition to focusing on improving resilience 
    to such catastrophes and easing the financial burden on homeowners.
\end{abstract}

\section{Introduction}
\label{sec:intro}
As the frequency and severity of weather catastrophe increase, it is shown to pose threats to homeowners' safety and finance.
It investigates their impact on property insurance pricing, emphasizing the catastrophic consequences for homeowners and insurers. 
Managing insurance premiums, deductibles, and coverage limitations becomes challenging for homeowners, while insurers face massive 
financial risks.

Unlike prior research \cite{hurricaneeco}, our study dives deeper into the dynamics of frequency and severity in property insurance 
pricing, seeking effective ways to alleviate the financial burden during major disasters. Drawing data from reliable sources,
the Insurance Information Agency, National Insurance Centers for Environmental Information, and Aon's 2023 Weather, Climate, and 
Disaster Investigation Report — where we will use statistical analysis and data science techniques. The objective is to enhance 
the well-being of homeowners and ensure the stability in the property insurance market against the increasing weather-related risks.

\section{Data}
\label{sec:data}
In order to understand the impact of weather catastrophe on property insurance pricing, we will first look at a dataset that 
contains important information about the frequency, financial cost, and some other parameters. Our data sources are drawn from 
National Centers for Environmental Information (NCEI)\cite{ncei}, the Insurance Information Institute (III)\cite{iii}, Aon\cite{aon} 
and the National Association of Insurance Commissioners (NAIC)\cite{naic}, which are all well known for their reliability in 
documenting catastrophic events and their associated economic and insured losses.

@@

The figure \ref{fig:disaster_threats} shows the cumulative billion-dollar disaster costs in the United States from 1980-2023. It 
illustrates how annual disaster costs accumulate through the year, with highlighting the key disasters of the year. The table 
also provides context on total disaster costs over time.

@@

The figure \ref{fig:regional_disasters} shows the number of billion-dollar disaster events by specific perils in the United States 
from 1980-2023.

@@

The chart \ref{fig:10economic_loss} showing the top ten most expensive cyclone economic losses from 1900 to 2022, highlighting 
the rising risks and costs that property insurers are facing. Specifically, the chart focuses on vulnerable coastal and island regions,
which are known to be cyclone, hurricane, and flooding hazardous zones. The chart \ref{fig:10global_loss} presents the top ten most expensive global insured catastrophe losses from 1900 to 2022, highlighting 
the rising risks and costs that property insurers are facing.

@@

@@

\begin{equation}
    \label{eq:freq}
    \text{Frequency} = \frac{\text{Number of Loss Events}}{\text{Exposure}}
\end{equation}

The frequency equation \ref{eq:freq} calculates the frequency by dividing the number of loss events by the exposure.

\begin{equation}
    \label{eq:sev}
    \text{Severity} = \frac{\text{Total Loss}}{\text{Number of Loss Events}}
\end{equation}

The severity equation \ref{eq:sev} calculates the severity by dividing the total loss by the number of loss events.

\begin{equation}
    \label{eq:premium}
    \text{Premium} = \lambda \cdot \text{Severity} \cdot \text{Exposure}
\end{equation}

The premium equation \ref{eq:premium} calculates the estimated premium based on the frequency, severity, and exposure. This should be noted
to be significant for both homeowners and insurance companies as it displays the financial consequences.

Through these key equations, we hope to gain a thorough understanding between the relationship of weather-related disasters 
and property insurance pricing.
